\systemEnding: Bootstrapping Domain-Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-10-31 archive.tex
got rid of old graphics content; deanonymized
possibly final talk slides
version space archive progress

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\indicator}{\mathds{1}} %{{\rm I\kern-.3em E}}
\newcommand{\expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}

@@

\begin{icmlauthorlist}
\icmlauthor{Kevin Ellis}{MIT}
\icmlauthor{Lucas Morales}{MIT}
\icmlauthor{Matthias}{MIT}
\icmlauthor{Maxwell Nye}{MIT}
\icmlauthor{Armando Solar-Lezama}{MIT}
\icmlauthor{Joshua B. Tenenbaum}{MIT}
\end{icmlauthorlist}

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
%In the \system algorithm we infer the DSL from a collection of frontiers.
Intuitively, we want the algorithm to
look at  the frontiers and
generalize beyond them, 
both so the DSL can better express the current solutions,
and  also so that the DSL might expose new abstractions
which will later be used to
discover more programs.
 Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
\label{AIC}
  \end{align}
We induce a DSL by searching locally through the space of DSLs,
proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC} fails to increase.
The search moves work by introducing new
$\lambda$-expressions into the DSL.
We propose these new expressions by extracting fragments of
programs already in the frontiers (Tbl.~\ref{fragmentExample}).
An important point here is that we are \emph{not} simply adding
subexpressions of programs to $\mathcal{D}$, as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316} and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

@@

\begin{comment}
  \begin{wrapfigure}{R}{0.6\textwidth}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}

@@

\begin{figure}  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{cl}\toprule
    \rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
    \begin{tabular}{l}
      \code{[4\, 7\, 6\, 9\, 0]$\to$5}\\
\code{[3\, 8\, 2\, 1\, 4]$\to$4}\\
\code{[2\, 2\, 9\, 4\, 8]$\to$3}\\
\code{[0\, 9\, 8\, 9\, 9\, 8]$\to$1}\\
\blueCode{$f(\ell) = $($f_0$ (car $\ell$))}\\\\

\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
\blueCode{$f(\ell) = $($f_4$ + $\ell$ 0)}\\\\

\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
\blueCode{$f(\ell) = $($f_5$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\

      \code{[5\, 2\, 9]}$\to$\code{[[2\, 9]\, [9]\, []]}\\
      \code{[3\, 8\, 1\, 3]}$\to$\code{[[8\, 1\, 3]\, [1\, 3]\, [3]\, []]}\\
      \blueCode{$f(\ell) = $($f_2$ empty? cdr cdr $\ell$)}
      \\

      \end{tabular}
    \\\midrule
    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{l}
      \greenCode{$f_0($x$)\,=\,$(+ x 1)}\\
      \hspace{\helpSize}($f_0$: \emph{increment})\\
      \greenCode{$f_1($x$)\,=\,$(- x 1)}\\
      \hspace{\helpSize}($f_1$: \emph{decrement})\\
        \popp{\code{$f_2($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_2$ (n x))))}}\\
        \hspace{\helpSize}($f_2$: \emph{unfold})\\
        \popp{\code{$f_3($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{($f_3$ ($f_1$ i) (cdr l))))}}\\
        \hspace{\helpSize}($f_3$: \emph{index})\\
        \popp{\code{$f_4($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
      \phantom{\code{$f_4($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_4$ (cdr l))))}}\\
        \hspace{\helpSize}($f_4$: \emph{fold})\\
        \greenCode{$f_5($f$,$l$)\,=\,$(if (empty? l) nil)}\\
        \phantom{\greenCode{$f_5($f$,$l$)\,=\,$(if }}\greenCode{(cons (f (car l)) ($f_5$ (cdr l)))}\\
        \hspace{\helpSize}($f_5$: \emph{map})

@@

      \end{tabular}
    %% : (lambda (#(#(lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) $3 ($4 ($5 $0) ($1 (cdr $0))))))))))) (lambda (car $0))) (lambda (lambda (+ $0 $1))) 0 $0))
    %% (lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) ($1 ($3 $0)) ($4 $0) ($5 $0)))))))))
    %% (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (#(lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) empty (cons ($3 (car $0)) ($1 (cdr $0))))))))) (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq? $0 $3)))))))
    \\\bottomrule 
    \end{tabular}
  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
  \end{figure}

@@

\bibliography{main}
\bibliographystyle{icml2018}
\end{document}

>>>>

\usepackage{verbatim}
 \usepackage{booktabs}
 % For algorithms
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{algpseudocode}% http://ctan.org/pkg/algorithmicx
\usepackage{tikz}
\usepackage{circuitikz}
\usetikzlibrary{fit,bayesnet}
%\usetikzlibrary{arrows.meta}
%\usetikzlibrary{positioning}
%\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.pathmorphing}
\tikzset{squiggle/.style={decorate, decoration={snake,amplitude=.4mm}}}
\usepackage{dsfont}
\usepackage{amsmath}

@@

\newcommand{\Expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\indicator}{\mathds{1}} %{{\rm I\kern-.3em E}}
\newcommand{\expect}{\mathds{E}} %{{\rm I\kern-.3em E}}
\newcommand{\probability}{\mathds{P}} %{{\rm I\kern-.3em P}}
\newcommand{\shift}[1]{\uparrow^{#1}}
\newcommand{\substitute}[2]{[\$ #1 \mapsto #2]}
\newcommand{\reduce}{\longrightarrow}
\newcommand{\manyReduce}{\longrightarrow^*}

\newtheorem{definition}{Definition}

@@

\begin{icmlauthorlist}
\icmlauthor{Kevin Ellis}{MIT}
\icmlauthor{Lucas Morales}{MIT}
\icmlauthor{Mathias}{MIT}
\icmlauthor{Maxwell Nye}{MIT}
\icmlauthor{Armando Solar-Lezama}{MIT}
\icmlauthor{Joshua B. Tenenbaum}{MIT}
\end{icmlauthorlist}

@@

The purpose of the DSL is to
offer a set of abstractions
that allow an agent to easily express solutions to the tasks at hand.
Intuitively, we want the algorithm to look at the programs found
during waking and generalize beyond them, both so the DSL can better
express the current solutions, and also so that the DSL might expose
new abstractions which will later be used to discover more programs.

Formally, we want the DSL maximizing $\int \lowerBound\;\mathrm{d}\theta$ (Sec.~\ref{mathematicalFraming}).
We replace this marginal with an AIC approximation, giving the following objective for DSL induction:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\probability[p|\mathcal{D},\theta]\Bigg)
\label{AIC}
\end{align}
At a high level, our approach is to search locally through the space
of DSLs, proposing small changes to $\mathcal{D}$ until Eq.~\ref{AIC}
fails to increase.  These small changes consist of introducing new
candidate $\lambda$-expressions into the DSL.

However, there is a snag with this simple approach:
whenever we add a new expression $e$ to the DSL,
the programs found during waking ($\left\{\mathcal{F}_x \right\}$)
are not written in terms of $e$.
Concretely, imagine we wanted to discover a new DSL procedure for doubling numbers,
after having found the programs
\code{(cons (+ 9 9) nil)} and \code{($\lambda$ (x) (+ (car x) (car x)))}.
As human programmers,
we can look at these pieces of code and recognize that,
if we define a new procedure called \code{double},
defined as \code{($\lambda$ (x) (+ x x))},
then we can rewrite the original programs as
\code{(cons (double 9) nil)} and \code{($\lambda$ (x) (double (car x)))}.
This process is a kind of refactoring
where a new subroutine is defined (\code{double})
and the old programs rewritten
in terms of the new subroutine.
Figure~\ref{mapFactor}
diagrams this refactoring
process for a more complicated
setting,
where the agent must rediscover the higher-order function \code{map}
starting from the basics of Lisp and the Y-combinator.

Formally, we want a DSL
which puts high probability on
any refactoring of the programs found during waking, maximizing:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\&+\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\manyReduce p}\probability[p'|\mathcal{D},\theta]\Bigg)
\label{factorObjective}
\end{align}
where $p'\manyReduce p$
is the standard notation for ``expression $p'$ evaluates to $p$ by the standard rules of $\lambda$-calculus''~\cite{pierce}.

Equation~\ref{factorObjective} captures the idea that we want
to add new components to the DSL while jointly refactoring our old
programs in terms of these new components.  But
this joint inference problem is intractable, because there are
infinitely many ways of refactoring a program.
We combine two ideas to make this refactoring process tractable:

\textbf{Idea 1:} Limit the degree to which
a piece of code can be refactored.
Instead of considering every refactoring,
bound the number of $\lambda$-calculus evaluation steps
separating a refactoring from its original program.
Formally,
we define the set of $n$-step refactorings as:
\begin{equation}
  R_n(p) = \left\{p'\;:\;p'\underbrace{\reduce p''\reduce\cdots\reduce}_{\text{$\leq n$ times}} p \right\}
\end{equation}
where $p_1\reduce p_2$ is the standard notation for ``$p_1$ rewrites to $p_2$
in one step according to the rules of $\lambda$-calculus''~\cite{pierce}.
For example,
\begin{align*}
  \code{((lambda (x) (x x)) (lambda (y) y))}\reduce&\\
  \code{((lambda (y) y) (lambda (y) y))}\reduce&\\
  \code{(lambda (y) y)}&
\end{align*}
Returning to Equation~\ref{factorObjective},
this approximation gives the following objective:
\begin{align}
\nonumber      \log \probability[\mathcal{D}]& + 
\argmax_{\theta}\Bigg(\log \probability[\theta|\mathcal{D}] - \|\theta\|_0 \\+&\sum_{x\in X}\log \sum_{p\in \mathcal{F}_x}\probability[x|p]\max_{p'\in R_n(p)}\probability[p'|\mathcal{D},\theta]\Bigg)
\label{limitedObjective}
\end{align}
In practice, setting the number of refactoring steps $n$ to 3 suffices
to give a competent DSL learning algorithm.  Although the number of
refactorings is now finite, it is still prohibitively large, and in
fact grows exponentially quickly both as a function of $n$ and as a
function of the size of the program being refactored.  Next, we show
how to tame this exponential explosion.

\textbf{Idea 2:} Rather than explicitly enumerate every refactoring,
we developed a technique for compactly representing and efficiently
calculating the set of all refactorings.  The key idea is to represent
the set of refactorings using a \emph{version
  space}~\cite{lau2001programming,mitchell1977version,polozov2015flashmeta}.
A version space is a tree-shaped data structure
that compactly represents a large set of programs.
We first define version spaces over $\lambda$-calculus expressions,
and then give a dynamic program for constructing a version space
that represents the set of possible refactorings (i.e., $R_n(p)$).

A version space is either:
\begin{itemize}
\item A deBuijn index: written $\$i$, where $i$ is a natural number
\item An abstraction: written $\lambda v$, where $v$ is a version space
\item A  application: written $(f \;x)$, where both  $f$ and $x$ are version spaces
\item A  union: $\uplus V$, where $V$ is a set of version spaces
\item The empty set, $\varnothing$
\item The set of all $\lambda$-calculus expressions, $\Lambda$
\end{itemize}
The purpose of a version space to compactly represent a set of programs.
We refer to this set as the \textbf{extension} of the version space:
\begin{definition}
  The \textbf{extension} of a version space $v$ is written $\denotation{v}$
  and is defined recursively as:
  \begin{align*}
    \denotation{\$i}& = \left\{\$i \right\}\\
    \denotation{\lambda v}& = \left\{\lambda e : e\in \denotation{v} \right\}\\
    \denotation{(v_1\; v_2)}& = \left\{(e_1\;e_2) : e_1\in \denotation{v_1},\;e_2\in \denotation{v_2} \right\}\\
    \denotation{\uplus V}& = \left\{e : v\in V,\;e\in \denotation{v} \right\}\\
    \denotation{\varnothing}& = \varnothing\\
    \denotation{\Lambda}& = \Lambda
    \end{align*}
\end{definition}
Version spaces also support efficient membership checking, which we
write as $e\in v$, and which is equivalent to $e\in \denotation{v}$.
Important for our purposes, it is also efficient to refactor the
members of a version space's extension in terms of a new DSL.
We define $\textsc{refactor}(v|\mathcal{D})$ inductively as:
\begin{align*}
  \textsc{refactor}(v|\mathcal{D}) &= \begin{cases}
    e\text{, if $e\in \mathcal{D}$ and $e\in v$}\\
    \textsc{refactor}'(v|\mathcal{D})\text{, otherwise.}
  \end{cases}\\
  \textsc{refactor}'(e|\mathcal{D})& = e\text{, if $e$ is a leaf}\\
  \textsc{refactor}'(\lambda b|\mathcal{D}) &= \lambda \textsc{refactor}(b|\mathcal{D})\\
  \textsc{refactor}'(f\;x|\mathcal{D}) &= \textsc{refactor}(f|\mathcal{D})\;\textsc{refactor}(x|\mathcal{D})\\
  \textsc{refactor}'(\uplus V|\mathcal{D}) &= \argmin_{e\in \left\{\textsc{refactor}(v|\mathcal{D})\;:\;v\in V \right\}}\ell(e|\mathcal{D})
  \end{align*}
where $\ell(e|\mathcal{D})$ for program $e$ and DSL $\mathcal{D}$
is the size of the syntax tree of $e$,
when members of $\mathcal{D}$ are counted as having size 1.
Concretely, $  \textsc{refactor}(v|\mathcal{D})$ calculates $\argmin_{p\in \denotation{v}}\ell(p|\mathcal{D})$.

In Appendix~\ref{appendixVersion}, we describe a dynamic program for
efficiently constructing a version space containing every $n$-step
refactoring.
This dynamic program, which we call $I\beta^n(p)$,
satisfies $\denotation{I\beta^n(p)} =  R_n(p)$.
In other words,
this dynamic program
builds a data structure
that represents the entire set of
refactorings -- but without having to explicitly enumerate all of the refactorings.

With this machinery in hand, we now have all the pieces needed to
learn a DSL (Algorithm~\ref{grammarInductionAlgorithm})

    \begin{algorithm*}[H]
    %\setcounter{algorithm}{2} % there are two algorithms in the main paper.
       \caption{DSL Induction Algorithm}
       \label{grammarInductionAlgorithm}
       \begin{algorithmic}[1]
         \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
%         \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
         \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
         \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
         \WHILE {true}         
         \STATE $\forall p\in \bigcup_{x}\mathcal{F}_x: $ $v_p\gets I\beta_n(p)$ \Comment{Construct a version space for each program}
         \STATE Define $\text{refactor}(p|\mathcal{D}') = \argmin_{e\in \denotation{v_p}} \ell (e|\mathcal{D}')$ \Comment{Rewrites $p$ assuming $\mathcal{D}'$ is the DSL}
         \STATE Define $L(\mathcal{D}',\theta) =  \prod_x \sum_{p\in \mathcal{F}_x} \probability[x|p]\probability[\text{\textsc{refactor}}(p|\mathcal{D}')|\mathcal{D}',\theta]$ \Comment{Likelihood if $(\mathcal{D}',\theta)$ were the DSL}
         \STATE Define $\theta^*(\mathcal{D}') = \argmax_\theta \probability[\theta|\mathcal{D}'] L(\mathcal{D}',\theta)$ \Comment{MAP estimate of $\theta$}
         \STATE Define $\text{score}(\mathcal{D}') = \log \probability[\mathcal{D}'] + L(\mathcal{D}',\theta^*) - \|\theta\|_0$ \Comment{objective function}
         \STATE components $\gets$ $\left\{\textsc{refactor}(v|\mathcal{D})\;:\;\forall x, \forall p\in \mathcal{F}_x, \forall v\in \text{children}(v_p) \right\}$ \Comment{Propose many new DSL components}
         \STATE proposals $\gets$ $\left\{\mathcal{D}\cup\left\{c \right\}\;:\;\forall c\in \text{components} \right\}$ \Comment{Propose many new DSLs}
         \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in \text{proposals}}\text{score}(\mathcal{D}') $\Comment{Get highest scoring new DSL}
         \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$\Comment{No changes to DSL led to a better score}
         \STATE $\mathcal{D}\gets\mathcal{D}'$ \Comment{Found better DSL. Update DSL.}
         \STATE $\forall x\;:\;\mathcal{F}_x\gets\left\{\text{\textsc{refactor}}(p|\mathcal{D})\;:\; p\in \mathcal{F}_x\right\}$\Comment{Refactor frontiers in terms of new DSL}
         \ENDWHILE
       \end{algorithmic}
    \end{algorithm}

\begin{figure*}
  \centering\begin{tikzpicture}[every node/.style={inner sep=1,outer sep=0}]
    \node(p1)[draw] at (-1,0) {
      \begin{tabular}{l}
      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
      \texttt{ (cons (+ (car l) (car l))}\\
      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
    \end{tabular}
    };
    \node(r1)[draw,inner sep=0,outer sep=0] at ([yshift=-3cm]p1.south) {
      \begin{tabular}{l}
  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
\texttt{ ($\lambda$ (z) (+ z z)))}
      \end{tabular}
    };

        \node(p2)[draw] at ([xshift=4.5cm]p1.east) {
    \begin{tabular}{l}
      \texttt{(Y ($\lambda$ (r l) (if (nil? l) nil}\\
      \texttt{ (cons (- (car l) 1)}\\
      \phantom{\texttt{(cons }}\texttt{ (r (cdr l))))))}
    \end{tabular}

    };
    \node(r2)[draw] at ([yshift=-3cm]p2.south) {
\begin{tabular}{l}
  \texttt{(}\orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l)}}\\
\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{nil}}\\
\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{(cons (f (car l))}}\\
\phantom{(($\lambda$ (f) (Y ($\lambda$ (r l)}\orange{\texttt{ (r (cdr l)))))))}}\\
\texttt{ ($\lambda$ (z) (- z 1)))}
      \end{tabular}

    };

    \draw [->] (p1.south)  --(r1.north) node[fill=white,midway] {refactor};
    \draw [->] (p2.south)  --(r2.north) node[fill=white,midway] {refactor};

    \node[draw](m) at (3,-7) {
\begin{tabular}{l}
          \fbox{\textsc{map}} = \orange{\texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if (nil? l) nil}}\\
  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(cons (f (car l))}}\\
  \phantom{\texttt{\emph{map}} = \texttt{($\lambda$ (f) (Y ($\lambda$ (r l) (if }}\orange{\texttt{(r (cdr l))))))}}\\
  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z z)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
  \code{(}\fbox{\textsc{map}}\code{ ($\lambda$ (z) (+ z 1)))}\quad(program rewritten w/ new \fbox{\textsc{map}} primitive)\\
      \end{tabular}      
    };
    \draw [->](r1.south)--(m.north);
    \draw [->](r2.south)--(m.north);
    \node[fill=white] at ([yshift=0.5cm]m.north) {\textbf{Compress (MDL/Bayes objective)}};
    %% \visible<2-3>{
    %%   \node[fill=orange!50!white,align=center] at (r1) {\Huge $10^{14}$\\\Huge  refactorings };
    %%   \node[fill=orange!50!white,align=center] at (r2) {\Huge $10^{13}$\\\Huge  refactorings };
    %% }
    %% \visible<3>{
    %%   \node[fill=orange!50!white,align=center] at ([yshift=7cm]m.north) {\Huge version space: set of programs\\
    %%   \huge Lau 2003; Gulwani 2012};
    %%   }
    %% \visible<4>{
    %%   \node[fill=orange!50!white,align=center] at (r1) {\Huge $\leq 10^6$\\\Huge version spaces };
    %%   \node[fill=orange!50!white,align=center] at (r2) {\Huge $\leq 10^6$\\\Huge version spaces };
    %% }

       \node(t1)[draw] at ([yshift=1.5cm]p1.north) {\begin{tabular}{ll}
           \textbf{Task}:&\texttt{(1 2 3)$\to$(2 4 6)}\\
           &\texttt{(4 3 4)$\to$(8 6 8)}
       \end{tabular}};
       \draw [->] (t1.south)  --(p1.north) node[fill=white,midway] {program search};
       \node(t2)[draw] at ([yshift=1.5cm]p2.north) {\begin{tabular}{ll}
           \textbf{Task}:&\texttt{(1 2 3)$\to$(0 1 2)}\\
           &\texttt{(4 3 4)$\to$(3 2 3)}
       \end{tabular}};
       \draw [->] (t2.south)  --(p2.north) node[fill=white,midway] {program search};

  \end{tikzpicture}
  \caption{DSL learning as code refactoring. For each task we discover programs during waking, then refactor the code from those programs to expose common subprograms (highlighted in \orange{orange}). Common subprograms are incorporated into the DSL when they increase a Bayesian objective. Intuitively, these new DSL components best compress the programs found during waking.}\label{mapFactor}
  \end{figure*}

\begin{comment}

We propose these new expressions
by extracting fragments of programs already in the frontiers
(Tbl.~\ref{fragmentExample}).  An important point here is that we are
\emph{not} simply adding subexpressions of programs to $\mathcal{D}$,
as done in the EC algorithm~\cite{Dechter:2013:BLV:2540128.2540316}
and other prior work~\cite{DBLP:conf/ecai/LinDETM14}.  Instead, we are
extracting fragments that unify with programs in the frontiers.  This
idea of storing and reusing fragments of expressions comes from
Fragment Grammars~\cite{tim} and Tree-Substitution
Grammars~\cite{cohn2010inducing}, and is closely related to the idea
of antiunification~\cite{henderson2013cumulative}.

@@

  \end{comment}
\begin{comment}
  \begin{wrapfigure}{R}{0.6\textwidth}
  \begin{minipage}{0.6\textwidth}    
    \begin{algorithm}[H]
      \caption{The \system Algorithm}
      \label{mainAlgorithm}
      \begin{algorithmic}
        \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
        \STATE \textbf{Hyperparameters:} Enumeration timeout $T$
        %     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, recognition model $q(\cdot)$
        \STATE Initialize $\theta\gets \text{uniform}$ %, $q_0(\cdot ) = \theta_0$
        \FOR{$i=1$ {\bfseries to} $I$}
        %     \FOR{$x:\tau\in X$}
        \STATE  $\mathcal{F}^{\theta}_x\gets \{p| p\in
        \text{enum}(\mathcal{D},\theta,T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.1cm}(\textbf{Search})}
        \STATE $q\gets \text{train recognition model, maximizing }\mathcal{L}_{\text{RM}}$ \hspace{0.2cm}(\footnotesize{\textbf{Compile}})
        \STATE  $\mathcal{F}^{q}_x\gets\{p|p\in
        \text{enum}(\mathcal{D},q(x),T)\text{ if }\probability[x|p] > 0\}$
        \footnotesize{\hspace{0.15cm}\textbf{(Search)}}

@@

\begin{figure*}  \newcommand{\helpSize}{0.25cm}
  \begin{tabular}{cc}\toprule
    \normalsize \pop{Programs} \& Tasks&\normalsize \popp{DSL}\\\midrule

    %\rotatebox[origin=c]{90}{\normalsize \pop{Programs} \& Tasks}&
    \begin{tabular}{l}
\code{[2\, 1\, 4]$\to$[2\, 1\, 4\, 0]}\\
%\code{[9\, 4\, 8]$\to$[9\, 4\, 8\, 0]}\\
\code{[9\, 8]$\to$[9\, 8\, 0]}\\
\blueCode{$f(\ell) = $($f_2$ cons $\ell$ (cons 0 nil))}\\\\

\code{[2\, 5\, 6\, 0\, 6]$\to$19}\\
\code{[9\, 2\, 7\, 6\, 3]$\to$27}\\
\blueCode{$f(\ell) = $($f_2$ + $\ell$ 0)}\\\\

\code{[4\, 2\, 6\, 4]$\to$[8\, 4\, 12\, 8]}\\
%\code{[0\, 3\, 5\, 0\, 3\, 2]$\to$[0\, 6\, 10\, 0\, 6\, 4]}\\
\code{[2\, 3\, 0\, 7]$\to$[4\, 6\, 0\, 14]}\\
\blueCode{$f(\ell) = $($f_3$ ($\lambda$ (x) (+ x x)) $\ell$)}\\\\

      \code{[1\, 5\, 2\, 9]}$\to$\code{[1\, 2]}\\
      \code{[3\, 8\, 1\, 3\, 1\, 2]}$\to$\code{[3\, 1\, 1]}\\
      \blueCode{$f(\ell) = $($f_0$ empty? car}\\
      \phantom{\blueCode{$f(\ell) = $($f_0$}}\blueCode{($\lambda$ (l) (cdr (cdr l))) $\ell$)}
      \\

      \end{tabular}&

%    \rotatebox[origin=c]{90}{\normalsize \popp{DSL}}&
    \begin{tabular}{l}
        \popp{\code{$f_0($p$,$f$,$n$,$x$)\,=\,$(if (p x) nil}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(cons (f x) ($f_0$ (n x))))}}\\
        \hspace{\helpSize}($f_0$: \emph{unfold})\\
        \popp{\code{$f_1($i$,$l$)\,=\,$(if (= i 0) (car l)}}\\
      \phantom{\code{$f_1($f$,$l$,$x$)\,=\,$(if}}}\popp{\code{($f_1$ (- i 1) (cdr l))))}}\\
        \hspace{\helpSize}($f_1$: \emph{index})\\
        \popp{\code{$f_2($f$,$l$,$x$)\,=\,$(if (empty? l) x}}\\
      \phantom{\code{$f_2($f$,$l$,$x$)\,=\,$(if }}}\popp{\code{(f (car l) ($f_2$ (cdr l))))}}\\
        \hspace{\helpSize}($f_2$: \emph{fold})\\
        \greenCode{$f_3($f$,$l$)\,=\,$($f_2$ nil l ($\lambda$ (x a) (cons (f x) a)))}\\
        \hspace{\helpSize}($f_3$: \emph{map})\\
        \greenCode{$f_4(\ell)$\,=\,(if (empty? $\ell$) 0 (+ 1 ($f_4$ (cdr $\ell$)))))}\\
        \hspace{\helpSize}($f_4$: \emph{length})\\
        \greenCode{$f_5(n)$\,=\,($f_0$ (= $n$) ($\lambda$ (x) x) (+ 1) 0)}\\
        \hspace{\helpSize}($f_5$: \emph{range})

@@

      \end{tabular}
    %% : (lambda (#(#(lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) $3 ($4 ($5 $0) ($1 (cdr $0))))))))))) (lambda (car $0))) (lambda (lambda (+ $0 $1))) 0 $0))
    %% (lambda (lambda (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) ($1 ($3 $0)) ($4 $0) ($5 $0)))))))))
    %% (lambda (lambda (fix1 $0 (lambda (lambda (#(lambda (lambda (lambda (if $0 empty (cons $1 $2))))) (#(lambda (lambda (fix1 $0 (lambda (lambda (if (empty? $0) empty (cons ($3 (car $0)) ($1 (cdr $0))))))))) (lambda (#(+ 1) $0)) ($1 (#(lambda (- $0 1)) $0))) 0 (eq? $0 $3)))))))
    \\\bottomrule 
    \end{tabular}
  \caption{Bootstrapping a standard library of functional programming routines, starting from recursion along with primitive operations found in 1959 Lisp.}\label{learningFromScratch}
  \end{figure*}

@@

\bibliography{main}
\bibliographystyle{icml2018}

\appendix

\section{Appendix}

\subsection{Refactoring code with version spaces}\label{appendixVersion}
Observe that every $\lambda$-calculus expression is also a version space,
and that, if $e$ is an expression, then $\denotation{e} = \left\{e \right\}$.
